Generative AI for image reconstruction in Intensity Interferometry: a first attempt

### 2025-11-12 Phase_II/draft/response.txt

|  Reviewer:
|  Dear Editor,
|  Thank you for sending me this manuscript by Rai et al on image
|  reconstruction from intensity interferometry. It addresses an important
|  topic, and may after major revision be suitable for publication in AJ. At a
|  high level, the paper does not demonstrate that the GAN presented actually
|  recovers much of the information about the stars presented; or that it
|  would be able to reconstruct images of other objects or with different uv
|  sampling.
|  Kind regards,
|  The Reviewer
|  
|  Comments
|
|  - Rapidly rotating stars can be fit analytically to these data. It may be
|  worth considering an example of an astrophysical source that would be more
|  interesting to fit, where the data analysis problem is less analytically
|  tractable, such as for complex scenes (spotted stars, or disks, for
|  instance).

In this first attempt, is seems reasonable to consider only
gravity-darkened stars, for the following reasons.

@@

|  - There is quite an extensive lit review on the core ideas of intensity
|  interferometry and their corresponding equations, that could perhaps be
|  condensed.

The equations have been condensed now.

|  - In Figure 2 we have a single sparse sampling map. It appears to be used
|  in all of the other figures throughout the paper. How well does the network
|  perform with different uv sampling? If it's trained only on one, I would
|  expect very poorly - and it's extremely important to test this and think of
|  ways to extend to a more flexible sampling, because you really don't want
|  to have to train your image reconstruction model separately for every
|  single observation.

The (u,v) sampling for any given data set is fixed and known.
Training the network for a single observing run is a tolerable
computational cost.  Hence we train on the sampling used.

### 2025-11-12 Phase_II/draft/intro_reformatted.tex
edits

\section{Introduction}
The humanity instinctively feels a symbiotic relationship with the stars. One of the primary scientific projects of the humanity is to figure out what the stars are and how do they do what they do. The first obvious step in this project, beyond measuring their global parameters like diameter, mass, orbital and astrometric elements, is to obtain images of the stars with all details of the stellar surfaces. In case of the Sun, this is done routinely by observatories and sun-observation satellites. But such routine still remains a challenge even for the $\alpha$-centauri system, our nearest stellar neighbour. The objective is to achieve capability of high fidelity image reconstruction of distant stars. Two interferometry based techniques, namely, Michelson Interferometry (MI) or Intensity Interferometry (II) have emerged during the last century to address this objective. A discussion of the development of these two approaches and comparison of their respective merits and challenges can be found in \citep{Rai2025}. The work reported here presents the results of a first effort at applying a Conditional Generative Adversarial (neural) Network (C-GAN) to image reconstruction of a fast rotator using its simulated II observations.

>>>>

\section{Introduction}
Humans instinctively feels a relationship with the stars. One of the primary scientific projects of humanity is to figure out what the stars are and how do they do what they do. The first obvious step in this project, beyond measuring their global parameters like diameter, mass, orbital and astrometric elements, is to obtain images of the stars with all details of the stellar surfaces. In case of the Sun, this is done routinely by observatories and sun-observation satellites. But such routine still remains a challenge even for the $\alpha$-centauri system, our nearest stellar neighbour. The objective is to achieve capability of high fidelity image reconstruction of distant stars. Two interferometry based techniques, namely, Michelson Interferometry (MI) or Intensity Interferometry (II) have emerged during the last century to address this objective. A discussion of the development of these two approaches and comparison of their respective merits and challenges can be found in \citep{Rai2025}. The work reported here presents the results of a first effort at applying a Conditional Generative Adversarial (neural) Network (C-GAN) to image reconstruction of a fast rotator using its simulated II observations.
